---
title: "AssociationRule"
author: "Jason Antal"
date: "2024-08-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
library(arules)
library(arulesViz)
library(tidyverse)
setwd("C:/Users/jason/downloads") #CHANGE FILEPATH HERE
groceries <- read.transactions("groceries.txt", format = "basket", sep = ",")

```

Use the data on grocery purchases in groceries.txt and find some interesting association rules for these shopping baskets. The data file is a list of shopping baskets: one person's basket for each row, with multiple items per row separated by commas. Pick your own thresholds for lift and confidence; just be clear what these thresholds are and say why you picked them. Do your discovered item sets make sense? Present your discoveries in an interesting and visually appealing way.

Notes:

    This is an exercise in visual and numerical story-telling. Do be clear in your description of what you've done, but keep the focus on the data, the figures, and the insights your analysis has drawn from the data, rather than technical details.
    The data file is a list of baskets: one row per basket, with multiple items per row separated by commas. You'll have to cobble together your own code for processing this into the format expected by the "arules" package. This is not intrinsically all that hard, but it is the kind of data-wrangling wrinkle you'll encounter frequently on real problems, where your software package expects data in one format and the data comes in a different format. Figuring out how to bridge that gap is part of the assignment, and so we won't be giving tips on this front.


```{r}

print("In order to handle this data correctly, we will use the read.transactions method
      with format set to basket and separation set to a comma.")
setwd("C:/Users/jason/downloads")
groceries <- read.transactions("groceries.txt", format = "basket", sep = ",")

print("We'll mine frequent datasets using the apriori() method. I chose a support
of 0.005 to focus on items that appear in at least 0.5% of transactions, confidence 
of 0.3 to ensure a reasonable strength of association, and a lift greater than 
3 to identify rules where items appear together often enough to rule out random luck.")

rules <- apriori(groceries, 
                 parameter = list(support = 0.005, confidence = 0.3, minlen = 2),
                 control = list(verbose = FALSE))

# Filter rules by lift
rules <- subset(rules, lift > 3)

# Sort rules by lift
rules <- sort(rules, by = "lift", decreasing = TRUE)

# Inspect the top 10 rules
inspect(head(rules, 10))

# Plot top 20 rules
top_20_rules <- head(rules, 20)
plot(top_20_rules, method = "graph", engine = "htmlwidget")

cat("The top rules show strong associations between various types of vegetables and fruits. 
    For example, rule 1 suggests that customers who buy root vegetables are likely to also buy other vegetables.
    This makes sense as people buying fresh produce often buy a variety of items for meal preparation.")

# Create a scatter plot of support vs confidence
plot(rules, measure = c("support", "confidence"), shading = "lift")

# Create a parallel coordinates plot
plot(top_20_rules, method = "paracoord")

#Plot data by lift
ggplot(DATAFRAME(head(rules, 10)), aes(x = LHS, y = RHS)) +
  geom_tile(aes(fill = lift), color = "white") +
  geom_text(aes(label = round(lift, 2)), vjust = 1) +
  scale_fill_gradient(low = "yellow", high = "red") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Top 10 Association Rules",
       x = "Left Hand Side", y = "Right Hand Side",
       fill = "Lift")

cat("The heatmap visualizes the strength of association (lift) between different item pairs.
    Darker colors indicate stronger associations. We can see that certain vegetable combinations
    have particularly high lift values, suggesting they are often purchased together.")

```


