---
title: "ImageClassification"
author: "Jason Antal"
date: "2024-08-18"
output: pdf_document
---

```{r setup, include=FALSE #CHANGE FILEPATH AND RUN THIS BEFORE RUNNING BELOW}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
setwd("c:/Users/jason/downloads") #CHANGE FILEPATH HERE
social_marketing <- read_csv("social_marketing")
```

In this problem, you will train a neural network to classify satellite images. In the data/EuroSAT_RGB directory, you will find 11 subdirectories, each corresponding to a different class of land or land use: e.g. industrial, crops, rivers, forest, etc. Within each subdirectory, you will find examples in .jpg format of each type. (Thus the name of the directory in which the image lives is the class label.)

Your job is to set up a neural network that can classify the images as accurately as possible. Use an 80/20 train test split. Summarize your model and its accuracy in any way you see fit, but make you include at a minimum the following elements:

    overall test-set accuracy, measured however you think is appropriate
    show some of the example images from the test set, together with your model's predicted classes.
    a confusion matrix showing the performance of the model on the set test, i.e. a table that cross-tabulates each test set example by (actual class, predicted class).

I strongly recommend the use of PyTorch in a Jupyter notebook for this problem; look into PyTorch's ImageFolder data set class, which will streamline things considerably.

```{r cars}

```



```{r pressure, echo=FALSE}

```

