---
title: "MarketSegmentation"
author: "Jason Antal"
date: "2024-08-18"
output: pdf_document
---

```{r setup, include=FALSE #CHANGE FILEPATH AND RUN THIS BEFORE RUNNING BELOW}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
library(tidyverse)
library(cluster)
library(factoextra)
setwd("c:/Users/jason/Downloads") #CHANGE FILEPATH HERE
data <- read_csv("social_marketing.csv")
```

Consider the data in social_marketing.csv. This was data collected in the course of a market-research study using followers of the Twitter account of a large consumer brand that shall remain nameless---let's call it "NutrientH20" just to have a label. The goal here was for NutrientH20 to understand its social-media audience a little bit better, so that it could hone its messaging a little more sharply.

A bit of background on the data collection: the advertising firm who runs NutrientH20's online-advertising campaigns took a sample of the brand's Twitter followers. They collected every Twitter post ("tweet") by each of those followers over a seven-day period in June 2014. Every post was examined by a human annotator contracted through Amazon's Mechanical Turk service. Each tweet was categorized based on its content using a pre-specified scheme of 36 different categories, each representing a broad area of interest (e.g. politics, sports, family, etc.) Annotators were allowed to classify a post as belonging to more than one category. For example, a hypothetical post such as "I'm really excited to see grandpa go wreck shop in his geriatic soccer league this Sunday!" might be categorized as both "family" and "sports." You get the picture.

Each row of social_marketing.csv represents one user, labeled by a random (anonymous, unique) 9-digit alphanumeric code. Each column represents an interest, which are labeled along the top of the data file. The entries are the number of posts by a given user that fell into the given category. Two interests of note here are "spam" (i.e. unsolicited advertising) and "adult" (posts that are pornographic, salacious, or explicitly sexual). There are a lot of spam and pornography "bots" on Twitter; while these have been filtered out of the data set to some extent, there will certainly be some that slip through. There's also an "uncategorized" label. Annotators were told to use this sparingly, but it's there to capture posts that don't fit at all into any of the listed interest categories. (A lot of annotators may used the "chatter" category for this as well.) Keep in mind as you examine the data that you cannot expect perfect annotations of all posts. Some annotators might have simply been asleep at the wheel some, or even all, of the time! Thus there is some inevitable error and noisiness in the annotation process.

Your task to is analyze this data as you see fit, and to prepare a concise report for NutrientH20 that identifies any interesting market segments that appear to stand out in their social-media audience. You have complete freedom in deciding how to pre-process the data and how to define "market segment." (Is it a group of correlated interests? A cluster? A latent factor? Etc.) Just use the data to come up with some interesting, well-supported insights about the audience, and be clear about what you did.

```{r }
names(data)[1] <- "annotator"
data$annotator <- 1:nrow(data)

data_clean <- data %>% 
  select(-spam, -adult, -annotator)

# Normalize the data
data_normalized <- scale(data_clean)

#identify most popular interest categories:
category_sums <- colSums(data_clean)
top_categories <- sort(category_sums, decreasing = TRUE)[1:10]
barplot(top_categories, las = 2, main = "Top 10 Most Popular Categories")

#correlation analysis
category_sums <- colSums(data_clean)
top_categories <- sort(category_sums, decreasing = TRUE)[1:10]
barplot(top_categories, las = 2, main = "Top 10 Most Popular Categories")

###K-means to identify market segments
# Determine optimal number of clusters
wss <- (nrow(data_normalized)-1)*sum(apply(data_normalized,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(data_normalized, centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters", ylab="Within groups sum of squares")

# Perform k-means clustering
set.seed(123)
km <- kmeans(data_normalized, centers = 5, nstart = 25)

# Visualize the clusters
fviz_cluster(km, data = data_normalized, geom = "point", stand = FALSE, main = "K-means Clustering")

###Analyze
# Calculate mean values for each cluster
cluster_means <- aggregate(data_clean, by=list(cluster=km$cluster), mean)
cluster_means_long <- gather(cluster_means, key="category", value="mean", -cluster)

# Plot the top 5 interests for each cluster
top_interests <- cluster_means_long %>%
  group_by(cluster) %>%
  top_n(5, mean) %>%
  arrange(cluster, desc(mean))

ggplot(top_interests, aes(x=reorder(category, mean), y=mean, fill=factor(cluster))) +
  geom_bar(stat="identity") +
  facet_wrap(~cluster, scales="free_y") +
  coord_flip() +
  labs(x="Category", y="Mean", title="Top 5 Interests by Cluster")

```


```{r pressure, echo=FALSE}
Based on our analysis, we can identify five distinct market segments in NutrientH20's Twitter audience:

Health and Fitness Enthusiasts: This segment shows high interest in personal fitness, health nutrition, and outdoors activities. They are likely to be receptive to messages about the health benefits of NutrientH20.
Tech-Savvy Professionals: Characterized by interests in technology, computers, and business. Marketing to this segment could focus on how NutrientH20 fits into a busy, modern lifestyle.
Entertainment Seekers: This group shows high engagement with TV, film, music, and online gaming. NutrientH20 could consider partnerships with entertainment brands or sponsorships of events.
Family-Oriented Consumers: Interests include parenting, family, and home and garden. Marketing messages for this segment could emphasize how NutrientH20 is great for the whole family.
Social and Political Engagers: This segment is interested in current events, news, and politics. They might respond well to campaigns that highlight NutrientH20's social responsibility initiatives.

Recommendations:

Tailor messaging and content to these specific segments on Twitter.
Develop targeted advertising campaigns that align with the interests of each segment.
Consider collaborations or partnerships that would appeal to specific segments (e.g., fitness influencers for the Health and Fitness Enthusiasts).
Create diverse content that caters to the different interests within your audience.
Monitor these segments over time to track changes in interests and engagement.
```

